---
title: 'Data wrangling:  Exercise 2'
output:
  html_notebook: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Daily precipitation has been collected for 56 weather stations in the Colorado Front Range and made available as a R list contained in the binary R file [FrontRange.RData](./data/FrontRange.RData). Use the command

```{r eval=FALSE}
load("./data/FrontRange.RData")
```

to load the data set (called `FR`) in R. 

The data set FR is a list with 4 components "precip", "time", "info" and "Stot". The components "precip" and "times" are also lists, whereas "info" is a data frame and "Stot" is a vector. 

(1) The list "precip" has 56 components, one for each station. `FR$precip[[i]]` thus has all the daily rainfalls recorded for the i-th station. 

(2) In a similar structure, the component "time" of FR hold the dates at which the rainfalls are recorded. 

(3) The component "info" is a data frame with various information on the stations (longitude, latitude, elevation, number of records, plus some other variables that are not so important for us here). 

(4) Finally, the component "Stot" is a vector that contains the mean summer precipitation for each station.

Explore the dataset. Answer the following questions:

a. Are the number of entries recorded by each station the same?

b. Find the station at the highest elevation and the station at the lowest elevation. Plot the daily rainfalls over time recorded by
these two stations. What do you think is happening with the data after 1980?

c. For each of the two stations identified above, what is the proportion of days with 10mm of rain or more?

d. Write a function to find for a given station, the date of the record highest amount of precipitation.

e. Do you see any relationship between precipitation and elevation? What about between air pressure in summer and precipitation?
