---
title: "Stats 406 Homework 1"
author: "Xi Chen"
output: word_document
---
```{r}
#1(a)
U1=runif(1000,min=0,max=1)
U2=runif(1000,min=0,max=1)
vecZ=sqrt((-2*log(U1)))*cos(2*pi*U2)
vecZ[1:10]

#1(b)
mean(vecZ<0)
mean(vecZ<1&vecZ>0)
mean(vecZ>1)
```
Comments: The proportion of the elements of vecZ that fall below 0 is 0.475; the proportion of the elements of vecZ that fall between 0 and 1 is 0.362; the proportion of the elements of vecZ that fall above is 0.163.  

```{r}
#2
x=0.2
n=20
k=0:20
sum(abs(sin(2*pi*k/n))*choose(n,k)*x^(k)*(1-x)^(n-k))
```
Comments: Bn(x)=0.8099159
```{r}
load("C:/Users/chenxi/Desktop/Academics/16 Fall/STATS 406/Homework/FrontRange.RData")

#3(a)
which.max(FR$info$elev)
which.min(FR$info$elev)

plot(FR$time[[36]],FR$precip[[36]],xlab="Time",ylab="Rainfalls",main = "Station 36 (Highest elevation)")

plot(FR$time[[27]],FR$precip[[27]],xlab="Time",ylab="Rainfalls",main = "Station 27 (Lowest elevation)")
```

Comments: The station at the highest elevation is Station 36 while the one at the lowest elevation is Station 27.

```{r}
#3(b)
S36_more=FR$precip[[36]][which(FR$precip[[36]]>=10)]

S27_more=FR$precip[[27]][which(FR$precip[[27]]>=10)]

length(S36_more)/length(FR$precip[[36]])

length(S27_more)/length(FR$precip[[27]])
```
Comments:For Station 36 which is at the highest elevation, the proportion of days with 10mm of rain or more is 0.262334.
For Station 27 which is at the lowest elevation, the proportion of days with 10mm of rain or more is 0.1132226. 

```{r}
#3(c)
FR$time[[10]][which.max(FR$precip[[10]])]
```
Comments: In Station 10, the date of the recored highest amount of precipitation is 1999.592. 