---
title: "Homework 0 template"
author: "GAO Zheng"
date: "September 1, 2017"
output: 
  pdf_document:
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Coinflip Trivia

Write formula as you would in LaTeX

$$
\mathbb P(HT|T) = p.
$$

# Exchangeability

## Use subsections for sub-questions

## Match the sub-question labels

This question has 2 sub-questions

# Gamma-Poisson

We strongly recommend following style guide for code chuncks. See, for example, [Advanced R](http://adv-r.had.co.nz/Style.html) [http://adv-r.had.co.nz/Style.html], or [Google's R stype guide](https://google.github.io/styleguide/Rguide.xml) [https://google.github.io/styleguide/Rguide.xml].

Always set seed in your simulation

```{r}
set.seed(123)
(N <- rpois(n = 1, lambda = 5))
```

# Gaussian Example

Always label axes in your plots

```{r fig.height = 4}
par(mar = c(4,4,1,1), las = 1)
x <- seq(-10, 10, length = 1000)
plot(x, sin(x), xlab = "x-values", ylab = "f(x)", 
     type = "l" , cex = 0.5, ylim = c(-1.2, 1.5))
lines(x, cos(x), lty = 4, lwd = 2)
abline(h = 0)
legend("topright", legend = c("y = sin(x)", "y = cos(x)"),
       lty = c(1, 4), bty = "n")
```

